---
title: 'Assignment #1 (formative)'
author: "Edimilson Torres de Oliveira Neto"
date: "2023-10-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1


My forked version of the assignment template repository can be found at https://github.com/editor210/472_assignment_template


## Exercise 2

For the second exercise I have chosen to work with the ChickWeight Dataset.
This Dataset presents 578 rows and 4 variables (Weight, time, Chick, Diet).

This dataset can be considered tidy since it fulfills the three rules for tidy data, that means:
1. Each Variable has its own column
2. Each observation has its own row
3. Each Value has its own cell

Looking at the table below it is possible to see that:
1. the variables "Weight", "Time", "Chick" and "Diet" are all contained in different columns
2. Each of the rows represent a unique weight given the chick, time, and diet
3. each value is separated in its own cell

```{r chicks, echo= FALSE}
head(ChickWeight)
```

One alternative to untidy the table would be to pivot the time variable into columns, maintaining the weight of the Chick as value.

This way the table breaks the first rule of the tidy data, as it is possible to see on the table below

```{r chicks, echo= FALSE}
library(tidyverse)
data <- ChickWeight

untidy_data <- pivot_wider(data, names_from = Time, values_from = weight)
head(untidy_data)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
