---
title: 'Assignment #1 (formative)'
author: "Edimilson Torres de Oliveira Neto"
date: "2023-10-02"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#importing the tidyverse library
library(tidyverse)
```

## Exercise 1


My forked version of the assignment template repository can be found at https://github.com/editor210/472_assignment_template


## Exercise 2

For the second exercise I have chosen to work with the ChickWeight dataset.
This dataset contains 578 rows and 4 variables (Weight, time, Chick, Diet).

This dataset can be considered tidy since it adheres to the three rules for tidy data, that means:
1. Each Variable has its own column
2. Each observation has its own row 
3. Each Value has its own cell

The table below shows the first six rows of the ChickWeight dataset. Looking at this table it is evident that:

1. the variables "Weight", "Time", "Chick" and "Diet" are all contained in different columns.
2. Each of the rows represent a unique weight given the chick, time, and diet.
3. each value is separated in its own cell.

```{r chicks, echo= FALSE}
#Checking the first six rows of the table
head(ChickWeight)
```

One alternative to untidy the table would be to pivot the time variable into columns, maintaining the weight of the Chick as a value, as it is shown in the next table.

```{r chicks_untidy, echo= FALSE}
#linking the dataset to a variable
db <- ChickWeight

#untidying the data by making it wider, linking the columns to the Times variable. The Values correpond to the weight values.
untidy_data <- pivot_wider(db, names_from = Time, values_from = weight)
head(untidy_data)
```


To get this new table it was necessary to transform the original dataset into a tibble and use pivot_wider to transform the time values into columns.

Now the observations in the table show the weight of the Chick on a specific Time, but it is not clear what the observation is without prior explanation.

In conclusion, this new table breaks the second rule of a tidy table because the "Time" and "Weight" variables are not allocated in separate columns


## Exercise 3

I chose the "Edgar Anderson's Iris Data" dataset for this exercise. This dataset gives the measurements in centimeters of sepal length and width and petal length and width for 50 flowers of 3 species.


```{r exercise_3, echo=TRUE}
#Importing the data
db <- iris

#checking the format of the data
head(db)

#Using the pipe operator to get the maximum value of the Sepal.Length variable
db %>% select(Sepal.Length) %>% max()

#grouping the data by species and getting the average value of all variables, by calling the group_by() and summarise() functions
summarized_data <- db %>% group_by(Species) %>% summarise(Mean_Petal.Length = mean(Petal.Length), Mean_Petal.Width = mean(Petal.Width), Mean_Sepal.Length = mean(Sepal.Length), Mean_Sepal.Width = mean(Sepal.Width))
print(summarized_data)

#Considering the irregular shape of sepals and petals, for this exercise I will consider the area as if they were rectangular, but that might not be applicable in real studies.
# Therefore, I use the summarized data to perform the transformation of the numeric variable by calculating the approximate mean area of sepals and petals by species.

transformed_data <- summarized_data %>% 
  group_by(Species) %>%
  summarise ("Mean Petal Area (cm²)" = Mean_Petal.Width * Mean_Petal.Length, "Mean Sepal Area(cm²)" = Mean_Sepal.Width * Mean_Sepal.Length)

#Checking the final format of the data
head(transformed_data)

```

After having made the above-mentioned transformations in the dataset, it is possible to  get some information that was not explicit in the initial dataset. For instance, the Setosa species presents Petals that are more than 15 times larger than the ones from the Versicolor flowers. However, when you check the sepal area, the setosa actually has, on average, a bigger Sepal than the versicolor species. The virginica species is also clearly the biggest flower measured in this dataset on average.

## Appendix: All code in this assignment

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE} 
# this chunk generates the complete code appendix. 
# eval=FALSE tells R not to run (``evaluate'') the code here (it was already run before).
```