---
title: 'Assignment #1 (formative)'
author: "Edimilson Torres de Oliveira Neto"
date: "2023-10-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#importing the tidyverse library
library(tidyverse)
```

## Exercise 1


My forked version of the assignment template repository can be found at https://github.com/editor210/472_assignment_template


## Exercise 2

For the second exercise I have chosen to work with the ChickWeight Dataset.
This Dataset presents 578 rows and 4 variables (Weight, time, Chick, Diet).

This dataset can be considered tidy since it fulfills the three rules for tidy data, that means:
1. Each Variable has its own column
2. Each observation has its own row
3. Each Value has its own cell

The table below shows the first six rows of the ChickWeight dataset. Looking at this table it is possible to see that:
1. the variables "Weight", "Time", "Chick" and "Diet" are all contained in different columns
2. Each of the rows represent a unique weight given the chick, time, and diet
3. each value is separated in its own cell

```{r chicks, echo= FALSE}
#Checking the first six rows of the table
head(ChickWeight)
```

One alternative to untidy the table would be to pivot the time variable into columns, maintaining the weight of the Chick as value, as it is shown in the next table.

```{r chicks_untidy, echo= FALSE}
#linking the dataset to a variable
db <- ChickWeight

#untidying the data by making it wider, linking the columns to the Times variable. The Values correpond to the weight values.
untidy_data <- pivot_wider(db, names_from = Time, values_from = weight)
head(untidy_data)
```


To get this new table it was necessary to transform the original dataset into a tibble and use pivot_wider to transform the time values into columns.

Now the observations on the table show the weight of the Chick on a specific Time, but it is not clear what the observation is without prior explanation.

In conclusion, this new table breaks the second rule of the tidy table because the "Time" and "Weight" variables are not allocated in separate columns


## Exercise 3


Choose a different built-in dataset within R and load it into memory.

Your task is to transform this data frame and demonstrate the use of at least the following three components:

A piping operator: either %>% or |>

A call to the group_by() and summarise() functions

A transformation of a numeric variable after having called summarise()


```{r exercise_3, echo=TRUE}
#Importing the data
db <- iris

#checking the format of the data
head(db)

#Using pipe to get the maximum value of the Sepal.Length variable
db %>% select(Sepal.Length) %>% max()

#grouping the data by species and getting the average value of all variables
db %>% group_by(Species) %>% summarise(Mean_Petal.Length = mean(Petal.Length), Mean_Petal.Width = mean(Petal.Width), Mean_Sepal.Length = mean(Sepal.Length), Mean_Sepal.Width = mean(Sepal.Width))

#Transforming the numeric variable after checking the summarizes values by calculating the mean area of sepals and petals by species.
db %>% group_by(Species) %>%
  summarise(Mean_Petal.Length = mean(Petal.Length), Mean_Petal.Width = mean(Petal.Width), Mean_Sepal.Length = mean(Sepal.Length), Mean_Sepal.Width = mean(Sepal.Width)) %>% 
  group_by(Species) %>%
  summarise ("Mean Petal Area" = Mean_Petal.Width * Mean_Petal.Length, "Mean Sepal Area" = Mean_Sepal.Width * Mean_Sepal.Length)


```
## Appendix: All code in this assignment

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE} 
# this chunk generates the complete code appendix. 
# eval=FALSE tells R not to run (``evaluate'') the code here (it was already run before).
```